MSD panels 1, 2 and 3 HVE Supernatant

### 2016-09-22 MSD_panels_1_2_3_HVE_sup_analysis.Rmd
get logFC ratio by doing difference of log mock subtracted for 24hr and 3 hr (bcause log(a) -log(b) = log(a/b)

---
title: "MSD panels 1, 2 and 3 HVE Supernatant"
output: github_document
---

@@

```{r filter for 1:2, echo=FALSE, message = FALSE, warning = FALSE}
### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection.

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow %>%
  filter(Sample_Dilution =="1:2", Detection.Range != "Above Fit Curve Range")

@@

HVE_all_panels_noStds_AdjBelow_short <- rbind(HVE_all_panels_noStds_AdjBelow_short,IL1RA_4_T3,IL1RA_2_T2)

@@

#Result is a df where the samples with dischordant replicates in terms of Detection.Range. The samples with "1" in the Number_of_Reps colimn are the ones where I want to use the data from the rep that was in/above range for the avg, NOT the average between the in-range rep and the below-range rep. Recall that I assigned all below range/curve reps a value of 0.01. If both reps were below, ok to use calculate the average of them.

@@

mock <- avgConc %>%
  filter(Virus == "M")

@@

no_mock_avgConc <-
  avgConc %>%
  filter(Virus !="M")

@@

#check for NA's (places where there was a mock sample but no corresponding V1 or V2)
whichNAs<-mergeMock[which(is.na(mergeMock$Sample.x)),]

#no NA's!

@@

#When the experimental and mock samples were both "below", we get zero in the minusMockEstConc column, and that won't show up on a log scale. So I'm setting all zeros in the minusMockEstConc column to 0.01.

#Also, in cases where the Mock values are > Virus values, we would get a negative number for the result and not be able to plot it on a log scale. So I will take the log of each, then subtract, then plot on a linear scale, so I can have the negatives there.

#Recall that log(a/b) = log(a)-log(b)
#so if Mock = 100 and V = 10, log 100/10 = log (0.1) = -1
# and log(100)-log(10) = 1-2 = -1

mergeMock <- mergeMock %>%
  mutate(logMock = log10(avgEstConcMock))%>%
  mutate(logVirus = log10(avgEstConcVirus))%>%
  mutate(MockSubtracted = logVirus - logMock)

@@

## Plots of Analyte Concentrations by Donor, Timepoint and Condition

## Caveats

@@

```{r  individual analyte plots, echo=FALSE,fig.width=8, fig.height = 15, message = FALSE, warning = FALSE}

ggplot(mergeMock, aes(x = Timepoint, y = MockSubtracted))+
   geom_jitter(aes(color = Donor), alpha = 0.75, size = 2.5, width = 0.2, height = 0.2)+
   labs( y = "log 10 Estimated Stock Concentration pg/ml")+
   facet_grid(Assay~Virus, scales = "free_x")+
   theme(axis.text.x = element_text(size = 11))+
  ggtitle("Analyte Concentrations ")

@@

>>>>

---
title: "MSD panels 1, 2 and 3 HVE Supernatant"
output: html_document
---

@@

```{r filter for 1:2 MOSTLY, echo=FALSE, message = FALSE, warning = FALSE}
### NOTE ### I'm going to filter for just the 1:2 dilution samples here since they gave more in-detection. its ca''ed short because theres only 1:2 data

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow %>%
  filter(Sample_Dilution =="1:2")

@@

HVE_all_panels_noStds_AdjBelow_short <- HVE_all_panels_noStds_AdjBelow_short %>%
  filter(!(Assay == "IL1-RA" & Donor == 2 & 
             Timepoint == "T2" & Sample_Dilution=="1:2")) %>% 
  filter(!(Assay == "IL1-RA" & Donor == 4 & 
             Timepoint == "T3" & Sample_Dilution=="1:2"))

HVE_all_panels_noStds_AdjBelow_short <- rbind(
  HVE_all_panels_noStds_AdjBelow_short, IL1RA_4_T3, IL1RA_2_T2)

```{r make adjustments for nice plotting, echo=FALSE, message = FALSE, warning = FALSE}

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow_short %>%
  mutate(Timepoint = ifelse(Timepoint == "T1", "3",
                            ifelse(Timepoint=="T2","8","24")))

#factor to get order right
HVE_all_panels_noStds_AdjBelow_short$Timepoint<-factor(HVE_all_panels_noStds_AdjBelow_short$Timepoint,levels = c("3","8","24"))

HVE_all_panels_noStds_AdjBelow_short<-HVE_all_panels_noStds_AdjBelow_short%>%
  mutate(Virus = ifelse(Virus == "V1", "SD90",
                        ifelse(Virus == "V2", "186","Mock")))

HVE_all_panels_noStds_AdjBelow_short$Virus<-factor(HVE_all_panels_noStds_AdjBelow_short$Virus,levels = c("Mock","SD90","186"))

@@

#Result is a df where the samples with discordant replicates in terms of Detection.Range. The samples with "1" in the Number_of_Reps colimn are the ones where I want to use the data from the rep that was in/above range for the avg, NOT the average between the in-range rep and the below-range rep. Recall that I assigned all below range/curve reps a value of 0.01. If both reps were below, ok to use calculate the average of them.

@@

mock <- avgConc %>%
  filter(Virus == "Mock")

@@

no_mock_avgConc <-
  avgConc %>%
  filter(Virus !="Mock")

@@

#check for NA's (places where there was a mock sample but no corresponding V1 or V2)
whichNAs<-mergeMock[which(is.na(mergeMock$Sample.x)),]
#no NA's!

@@

#do the subtraction, with the result being the difference between the log10 of virus - log 10 Mock
mergeMock <- mergeMock %>%
  mutate(logMockSubtracted = log10(avgEstConcVirus) - log10(avgEstConcMock))

@@

Samples are multiplied by their dilution factor to get estimated concentrations.

@@

```{r setup for individual analyte HEATMAP, echo=FALSE}

#I want to make a heatmap that has the following data

#NO 8hr timepoint data

#tiles represent foldchange of Mock subtracted 24hr data over mock-subtracted 3hr. So if 24hr>3hr, you get a ratio <1 and if 24hr <3hr you get a ratio >1

#Want gradient intensity to max out at FC of 10.

#Exclude 8hr data.

no8mergeMock <- mergeMock %>%
  filter(Timepoint != "8")

# Extract the 3hr data and select a few columns

timepoint3<- no8mergeMock %>%
  filter(Timepoint == "3")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

#extract 24 hr data and select a few columns
timepoint24<- no8mergeMock %>%
  filter(Timepoint =="24")%>%
  select(Donor, Assay, Sample, Virus, logMockSubtracted)

@@

```{r merge 24 and 3, check NAs, echo=FALSE}
#merge 24hr and 3hr data

merge24And3<- merge(timepoint24,timepoint3, by  = c("Donor","Assay","Virus"), all.x = TRUE, all.y = TRUE)

#There are NO rows where we have data for 24hr but NOT for 3hr.
which24no3<-merge24And3[which(is.na(merge24And3$Sample.y)),]

#there are NO rows where we have data for 3hr but not for 24hr
which3no24<-merge24And3[which(is.na(merge24And3$Sample.x)),]

#Fix up colnames      
names(merge24And3)[4:7]<- c("Sample.24", "logMockSubtracted.24","Sample.3", "logMockSubtracted.3")
```

```{r get ratio, check NAs, echo=FALSE}

#Do the subtraction to get a foldchange.

merge24And3 <- merge24And3 %>%
  mutate(logMockSub24_minus_logMockSub3 = logMockSubtracted.24-logMockSubtracted.3)
```

```{r setup logFC heatmap, check NAs, echo=FALSE}

#I want the scale to have max intensity a 10-fold change in either direction (log10(10)=1 on my scale and log10(0.1)= -1 on my scale) so I am going to set all values >=1 as 1 and all values <=-1 as -1.

#set all values >=1 as 1

merge24And3 <- merge24And3 %>%
  mutate(max10fold =
           ifelse(
             logMockSub24_minus_logMockSub3 >=1,
         1,
           ifelse(
             logMockSub24_minus_logMockSub3 <=-1,
         -1,logMockSub24_minus_logMockSub3)
           )
  )

```{r logFC heatmap, check NAs, echo=FALSE}

ggplot(merge24And3, aes(x = Donor, y = Assay))+
  geom_tile(aes(fill = max10fold), color = "gray")+
  scale_fill_gradient2(low = "green", mid = "white", high = "red", midpoint = 0)+
  facet_wrap(~Virus)+
  ggtitle("log fold change of concentrations at 24hrs over 3 hrs")
```

```{r trying pheatmap, check NAs, echo=FALSE}

#trying with pheatmap
# library(pheatmap)
# library(tidyr)
# 
# #trying with pheatmap since ggplot is funky looking and I want zero in the middle
# 
# #pheatmap need it as numeric matrix and can't facet so I'll do v186 separately from sd90
# 
# SD90heatmap<-merge24And3 %>%
#   filter(Virus == "SD90")%>%
#   select(Donor, Assay, max10fold)
# 
# 
# matrixSD90heatmapSpread<-SD90heatmap%>%
#   spread(key = Donor, value = max10fold)%>%
#   select(-Assay)%>%
#   as.matrix()
# 
# 
# rownames(matrixSD90heatmapSpread) <-SD90heatmapSpread$Assay
# 
# 
# pheatmap(matrixSD90heatmapSpread)
#

test = matrix(rnorm(200), 20, 10)
test[1:10, seq(1, 10, 2)] = test[1:10, seq(1, 10, 2)] + 3
test[11:20, seq(2, 10, 2)] = test[11:20, seq(2, 10, 2)] + 2
test[15:20, seq(2, 10, 2)] = test[15:20, seq(2, 10, 2)] + 4
colnames(test) = paste("Test", 1:10, sep = "")
rownames(test) = paste("Gene", 1:20, sep = "")

pheatmap(test)
